---
title: "Comparision of different runs"
author: "Shashank Tiwari"
date: "Updated<br> `r Sys.Date()`"
output: 
  rmarkdown::html_document: 
    theme: spacelab
    highlight: zenburn 
    code_folding: show 
    toc: true 
    toc_float: true
    smooth_scroll: true
    number_sections: false 
    self_contained: true 
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=T, message=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi=350,
  fig.height = 6,
  fig.width = 7.00787 #178mm
)  
knitr::opts_chunk$set(warning = FALSE) 
options(knitr.duplicate.label = "allow")
#Input
library(knitr)
library(png)
library(gridExtra)
library(ggplot2)
library(GenomicRanges)
library(chicane)
library(data.table)
library(dplyr)
library(stringr)
library(cowplot)
library(viridis)
library(slickR)
library(svglite)
library(patchwork)
library(magick)
library(DT)
library(kableExtra)
```


In our initial run of Chicane we observed that output contained almost 100% of bait to bait interactions. Another key thing we noticed was the number of unique baits and otherEnds. The output had far lesser number of baits present than compared to the input baits file. Apart from this it was noted that a great proportion of otherEnds were also classified as baits hence, when we looked at the lengths of the otherEnds we observed that Chicane has merged otherEnds and the length of the otherEnds was much higher than what we expected to see.

## Solutions

To overcome these issues we looked to rmove the overlapping regions from the input otherEnd and baits file as suggested by Dr. Syed. We observe that raw data has around 33k unique RNAs as baits, however if remove the baits which have an overlap with each other we are left with 22k unique baits. We are not sure if this approach is viable as we would lose critical information from those baits. It was also suggested that instead of usign the otherEnds from the raw data form Radicl-seq we use bins of 2kb for the entire genome for running chicane. Here we compare the outputs from these different approaches to understand what method is best suited for our use.

## The run
1. Initialy we use the otherEnd from the raw data file, the unique DNA tag rows. We remove the overlapping regions within this file using genomic ranges. Next we intersect this file with the baits file. Note that the baits file does contain overlappng regions.
2. Similar to the first run we define the otherEnds however this time we remove the overlapping regions from the baits file before we intersect it with the fragment file.
3. For this run we use the entire genome split into bins of 2kb as our fragments file. These bins are intersected with baits file containg overlapping regions.
4. Similar to the previous run we still use the 2kb bins however this time we use a baits file which has overlapping regions removed from itself.
5. We use 2kb bins for the other ends but the baits are filtered smartly, instead of merging the overlapping baits, we reduce the length of the bait by changing the end position of the baits.



We first examine the input baits file. The original baits contains the loaction of all the RNAs including the overlapping regions. It has a total of 60k entries, we reduce the overlapping regions and have a total of 41k baits in the input file. The RNAs which are dropped are the ones which are totaly contained inside another bait region.

```{r,eval=FALSE}
#Read the data
res_with_ovrlap <- fread("/home/master/shared_folder/radicl_seq/Data/Processed/graphs/chicane.results.txt")
res_wo_ovrlap <- fread("/home/master/shared_folder/radicl_seq/Data/Processed/graphs/chicane.results.wo.baitsoverlap.txt")
res_wo_ovrlap$target.len <- res_wo_ovrlap$target.end-res_wo_ovrlap$target.start
res_binned <- fread("/home/master/shared_folder/radicl_seq/Data/Processed/graphs/chicane.resuls.binned.txt")
res_binned$target.len <- res_binned$target.end-res_binned$target.start
res_binned_wo_ovrlap <-
  fread("/home/master/shared_folder/radicl_seq/Data/Processed/graphs/chicane.results.binned.wo.baits.txt")
res_binned_wo_ovrlap$target.len <- res_binned_wo_ovrlap$target.end-res_binned_wo_ovrlap$target.start
res_binned_2  <- fread("/home/master/shared_folder/radicl_seq/Data/Processed/graphs/chicane.results.binned.wo.baits_2.txt")
res_binned_2$target.len  <- res_binned_2[,target.end-target.start]
res_binned_2$type <- as.factor("binned_with_proper_baits")
res_with_ovrlap$type <- as.factor("raw_data")
res_wo_ovrlap$type <- as.factor("raw_data_wo_ovrlap")
res_binned$type <- as.factor("binned")
res_binned_wo_ovrlap$type <- as.factor("binned_wo_ovrlap")
merged_runs <- rbind(res_with_ovrlap,res_wo_ovrlap,res_binned,res_binned_wo_ovrlap,res_binned_2)
merged_runs$signif <- ifelse(merged_runs$q.value<0.05,"significant","non-significant")
merged_runs$bait.trans.count <- NULL
merged_runs$target.trans.count <- NULL
```
```{r,echo=FALSE}
load("merged.RData")
```




The first major thing we notice is that none of the outputs have a NA q.values as we had seen before.This indicates that the model is fit better when overlapping regions are removed

## Comparing number of unique baits in the output.

```{r,echo=FALSE}
Runs <- c("Raw data","Raw data(wo_ovrlap)","Binned","Binned(wo_ovrlap)")
Baits_Input <- c(33931,22093)
Baits_Output <- c(32955,21647,33060,21700)
Bait_to_Bait_percentage <- c(41.56,48.51,40.64,47.68)
Frags_input <- c(7435356,7423518,680094,668256)
Frags_output <- c(7373301,7384342,551739,540627)
NonbaitFrags_input <-c(7401425,7401425,658001,646163)
NonbaitFrags_output <- c(7352301,7352144,519619,519423)
dt <- data.table(Runs,Baits_Input,Baits_Output,Bait_to_Bait_percentage)
tabl <- data.table(Runs,Frags_input,Frags_output,NonbaitFrags_input,NonbaitFrags_output)
kable(dt,format = "html")
```

We can still see that a small fractions of baits are being filtered out from the input file. This percentage is not very high though.

Similarly we can check if the otherEnds are being filtered out as well.
```{r}
kable(tabl,format = "html")
```
Therefore we can see that chicane filters out fragments from the output. For the run with 2kb runs this is however expected as we are not including zero counts but we still see fragments being filtered in the other runs.



## Fragment Lengths

To check if chicane is still merging the otherEnds we look at the lengths of the otherEnds and compare it with the input file. We do so by looking at the otherEnds which are not classified as baits.

```{r}
ggplot(merged_runs[!merged_runs$target.id %in% merged_runs$bait.id & merged_runs$type %in% c('raw_data','raw_data_wo_ovrlap')]
       ,aes(x=signif,y=log10(target.len),fill=type))+
  geom_violin(trim = TRUE)+
  geom_boxplot(width=0.1,outlier.shape = NA)+
  theme_cowplot()+
  facet_wrap(~type)+
  scale_fill_viridis(discrete = T,alpha = 0.5)+
  theme(legend.position = "none")+
  ylab("log10(otherEnd Length)")+
  xlab("Type of run")+
  scale_y_continuous(limits = quantile(log10(merged_runs[!merged_runs$target.id %in%  merged_runs$bait.id]$target.len), c(0.1, 0.9)))
# ylim1 = boxplot.stats(log10(merged_runs[!merged_runs$target.id %in% merged_runs$bait.id]$target.len))$stats[c(1,5)]
# p1 = p0 + coord_cartesian(ylim = ylim1*1.05)
# plot(p1)
```
```{r,echo=FALSE,message=FALSE,results='hide'}
gc()
```


```{r} 
ggplot(merged_runs[!merged_runs$target.id %in% merged_runs$bait.id & !merged_runs$type %in% c('raw_data','raw_data_wo_ovrlap')]
       ,aes(x=signif,y=log10(target.len),fill=type))+
  geom_violin(trim = TRUE)+
  geom_boxplot(width=0.1,outlier.shape = NA)+
  theme_cowplot()+
  facet_wrap(~type)+
  scale_fill_viridis(discrete = T,alpha = 0.5)+
  theme(legend.position = "none")+
  ylab("log10(otherEnd Length)")+
  xlab("Type of run")+
  scale_y_continuous(limits = quantile(log10(merged_runs[!merged_runs$target.id %in% merged_runs$bait.id & !merged_runs$type %in% c('raw_data','raw_data_wo_ovrlap')]$target.len), c(0.1, 0.9)))
```

```{r,echo=FALSE,message=FALSE,results='hide'}
gc()
```

As we can see from the first plots that there is difference in the lenghts of the otherends for significant interactions, however for binned runs we cant get proper plots since most interactions are 2kb bins. We can check what percentage of interactions have otherEnds length greater than 2kb.

```{r}
lit = NULL
lit1 = NULL
for(i in unique(merged_runs$type)){
  lit[i] <- 100*length(merged_runs[merged_runs$type==i &merged_runs$target.len>1999 & merged_runs$signif=="non-significant"]$target.id)/
    length(merged_runs[merged_runs$type==i & merged_runs$signif=="non-significant"]$target.id)
  lit1[i] <- 100*length(merged_runs[merged_runs$type==i &merged_runs$target.len>1999 & merged_runs$signif=="significant"]$target.id)/
    length(merged_runs[merged_runs$type==i & merged_runs$signif=="significant"]$target.id)
}
tabl1 <- data.table(unique(merged_runs$type),lit,lit1)
colnames(tabl1) <- c("Runs","non-significant","Significant")
kable(tabl1,format = "html")
```

```{r,echo=FALSE,message=FALSE,results='hide'}
gc()
```

However we can remove bait-to-bait interactions from this list to see the overall results

```{r}
for(i in unique(merged_runs$type)){
   lit[i] <- 100*length(merged_runs[merged_runs$type==i &merged_runs$target.len>1999 & merged_runs$signif=="non-significant" & !merged_runs$target.id %in% merged_runs$bait.id]$target.id)/
     length(merged_runs[merged_runs$type==i & merged_runs$signif=="non-significant" & !merged_runs$target.id %in% merged_runs$bait.id]$target.id)
   lit1[i] <- 100*length(merged_runs[merged_runs$type==i &merged_runs$target.len>1999 & merged_runs$signif=="significant" & !merged_runs$target.id %in% merged_runs$bait.id]$target.id)/
     length(merged_runs[merged_runs$type==i & merged_runs$signif=="significant" & !merged_runs$target.id %in% merged_runs$bait.id]$target.id)
 }
tabl3 <- data.table(unique(merged_runs$type),lit,lit1)
colnames(tabl3) <- c("Runs","non-significant","Significant")
kable(tabl3,format = "html")
```

## Compare significant interactions



```{r}
Total <- c("0.094%","0.097%","0.132%","0.134%")
Cis <- c("17.52%","12.22%","16.56%","12.61%")
Trans <- c("82.47%","87.74%","83.43%","87.38%")
tabl2 <- data.table(Runs,Total,Cis,Trans)
kable(tabl2,format = "html")
```



For the significant interactions we can compare the distance of interactions. First we modify the data to suit the plots better

```{r}
signif <- merged_runs[merged_runs$q.value<0.05]
merged1<- signif[!is.na(distance),.(count2=1:count),names(merged_runs)]

```

```{r}
ggplot(merged1,aes(log10(distance),fill=type,colour=type))+
  geom_density(alpha=0.1)+
  theme_cowplot()+
  scale_fill_manual(name="Different runs",
                  values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
    scale_colour_manual(name="Different runs",
                        values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
    geom_vline(xintercept=log(300000,base=10), colour="grey") +
    annotate(x=log(300000,base=10),y=2,label="300Kb",vjust=2,geom="label")+
    xlab("Log 10 Distance from Interaction to RNA")
```


We can further normalise this distance with the RNA length and compare with our previous run

```{r}
merged1[,bait.length:=bait.end-bait.start]
merged1[,stnd_dist:=distance/bait.length]
ggplot(merged1,aes(log10(stnd_dist),fill=type,colour=type))+
  geom_density(alpha=0.1)+
  theme_cowplot()+
  scale_fill_manual(name="Different runs",
                  values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
    scale_colour_manual(name="Different runs",
                        values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
    geom_vline(xintercept=log(1,base=10), colour="grey") +
    geom_vline(xintercept=log(120,base=10), colour="grey") +
    annotate(x=log(3,base=10),y=1.8,label="RNA Length",vjust=2,geom="label")+
    annotate(x=log(400,base=10),y=1.8,label="120x RNA Length",vjust=2,
                geom="label")+
    xlab("Log 10 Distance from Interaction to RNA of RNA Length")

```


What we can conclude from these plots is that the results on are not very different based on the input fragment file, i.e. the fragment file from the raw data and the binned genome have almost identical distribution for cis interactions. A major difference can be seen if the overlapping regions from the baits file are removed. The peak shift very significantly without the overlapping regions. This is something we need to look further into and find a way to remove overlapping regions from the baits file without losing this critical information from those baits. 

We also should compare the trans significant interactions. Since we cannot look at distance for these interactions, we look at the counts distibution

```{r}
#merged2 <- rbind(signif_raw,signif_raw_wo,signif_bin,signif_bin_wo)
merged2_trans <- signif[is.na(distance)]
ggplot(merged2_trans,aes(log10(count),fill=type,colour=type))+
  geom_density(alpha=0.1)+
  theme_cowplot()+
  scale_fill_manual(name="Different runs",
                  values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
    scale_colour_manual(name="Different runs",
                        values=c("#69b3a2","blue","darkgoldenrod4","coral1","red"))+
  ggtitle("Distribution of counts")

```

```{r,eval=FALSE,echo=FALSE}
ggplot(merged_runs[is.na(merged_runs$distance)],aes(x=signif,y=count,fill=type))+
    geom_violin(trim = TRUE)+
    theme_cowplot()+
    ggtitle("Distribution of counts")+
    geom_boxplot(width=0.1,outlier.shape = NA)+
    theme_cowplot()+
    facet_wrap(~type)+
    scale_fill_viridis(discrete = T,alpha = 0.5)+
    theme(legend.position = "none")+
    ylab("counts")+
    xlab("Type of run")+
    scale_y_continuous(limits = quantile(merged_runs[is.na(merged_runs$distance)]$count, c(0.1, 0.9)))
```
```{r}
ggplot(merged2_trans,aes(x=signif,y=count,fill=type))+
    geom_violin(trim = TRUE)+
    theme_cowplot()+
    ggtitle("Distribution of counts")+
    geom_boxplot(width=0.1,outlier.shape = NA)+
    theme_cowplot()+
    facet_wrap(~type)+
    scale_fill_viridis(discrete = T,alpha = 0.5)+
    theme(legend.position = "none")+
    ylab("counts")+
    xlab("Type of run")+
    scale_y_continuous(limits = quantile(merged2_trans$count, c(0.1, 0.9)))
```

```{r,echo=FALSE,message=FALSE,results='hide'}
gc()
```